---
title: "cellmigRation"
output:
    html_document:
        toc: true
vignette: >
    %\VignetteIndexEntry{cellmigRation}
    %\VignetteEngine{knitr::rmarkdown}
    %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>", echo = TRUE, include = TRUE, eval = TRUE,
    message = FALSE,
    warning = FALSE,
    fig.align = "center", fig.keep = "last", fig.height = 5,
    fig.width = 9
)

Tim0 <- Sys.time()
set.seed(1234)

library(ggplot2)
library(dplyr)
library(cellmigRation)
library(kableExtra)
```

```{r set-options, echo=FALSE, cache=FALSE}
options(width = 14)
```


## Introduction

This vignette illustrates how to get started with **cellmigRation**, an R
library aimed at analyzing cell movements over time using multi-stack *tiff*
images of fluorescent cells.

The software includes two modules:

- **Module 1**: data import and pre-precessing. This module includes a series
of functions to import tiff images, remove noise/background and detect
cell/particles, (optional) automatically estimate optimal analytic
parameters, compute cell tracks (movements) and basic stats. The first module
is largely based on the FastTracks software written in Matlab by
*Brian DuChez* (FastTracks,
<https://www.mathworks.com/matlabcentral/fileexchange/60349-fasttracks>,
MATLAB Central File Exchange).

- **Module 2**: advanced analyses and visualization. The second module includes
a series of functions to compute advanced metrics/stats, exporting,
automatically built visualizations, and generate interactive/3D plots.

## Summary

This vignette guides the user through package installation, *tiff* file import,
cell tracking, and a series of downstream analyses.

- Package installation

- Module 1

    + Importing TIFF files

    + Optimizing Tracking Params

    + Tracking Cell Movements

    + Basic migration stats

    + Basic visualizations

    + Aggregate Cell Tracks

- Module 2

    + Import and Pre-process Cell Tracks

    + Plotting tracks (2D and 3D)

    + Deep Trajectory Analysis

    + Final Results

    + Principal Component Analysis (PCA) and Cell Clustering

### Notes and Acknowledgmenets

Damiano Fantini (Northwestern University, Chicago, IL, USA); Salim Ghannoum
(University of Oslo, Oslo, Norway)


### More resources

- An exhaustive vignette is available at: 
<https://www.data-pulse.com/projects/2020/cellmigRation/cellmigRation_v01.html>

- GitHub page: <https://github.com/ocbe-uio/cellmigRation>





## Installation

The package is currently available on Bioconductor. It can be installed using
the following command:

```{r}
try(
    BiocManager::install("cellmigRation", update = FALSE),
    silent = TRUE)
```


## cellmigRation Pipeline

#### Required libraries

For this demo, the following libraries have to be loaded.

```{r}
library(cellmigRation)
library(dplyr)
library(ggplot2)
library(kableExtra)
```

### Module 1

#### Importing TIFF files

In this vignette, we are going to analyze three images with the aim of
illustrating the functions included in *cellmigRation*. 
The original TIFF files are available at the following URLs:

- <https://www.data-pulse.com/projects/2020/cellmigRation/ctrl_001.tif>

- <https://www.data-pulse.com/projects/2020/cellmigRation/ctrl_002.tif>

- <https://www.data-pulse.com/projects/2020/cellmigRation/drug_001.tif>


**Note**. TIFF files can be imported using the `LoadTiff()` function. 
This function includes a series of (optional) arguments to attach 
meta-information to a TIFF image, for example the `experiment` and 
`condition` arguments. Imported numeric images are stored as 
a *trackedCells*-class object. 

Three sample `trackedCells` objects (imported from the corresponding TIFF files)
are available as a list in the `cellmigRation` package
(`ThreeConditions` object). These will be used for illustrating the functions
of our package in this vignette.


```{r echo=TRUE, include=TRUE, eval=TRUE, results='markup'}
# load data
data(ThreeConditions)

# An S4 trackedCells object
ThreeConditions[[1]]
```



#### Optimizing Tracking Params

This is an optional yet recommended step. Detecting fluorescent cells requires
defining a series of parameters to maximize signal to noise ratio. Specifically,

- **diameter**: size corresponding to the largest diameter of a cell (expressed
in pixels). Ideally, we want to set this parameter to a value large enough to
capture all cells (even the large ones), but small enough to exclude
aggregates or large background particles (artifacts, bubbles)

- **lnoise**: size corresponding to the smalles diameter of a cell (expressed
in pixels). Ideally, we want to set this parameter to a value small enough to
capture all cells (even the small ones), but large enough to exclude small
background particles (artifacts, debris)

- **threshold**: signal level used as background threshold. Signal smaller than
threshold is set to zero

If the values of these arguments are known, you can skip this step.
Alternatively, if you want to test a specific range of these values, you can
run `OptimizeParams()` manually specifying the ranges to be tested. By default,
the function determines automatically a reasonable range of values to be tested
for each argument based on the empirical distribution of signal and sizes of
particles detected in the frame with median signal from a TIFF stack. This
operation supports parallelization.

**Note**: the user may request to visualize a plot. The output plot shows how
many cells were detected for each combination of parameter values. By default,
the *pick #1* is selected for the downstream steps.

```{r echo=TRUE, include=TRUE, eval=TRUE, fig.height=7.8}
# Optimize parameters using 2 cores
x1 <- OptimizeParams(
    ThreeConditions$ctrl01, threads = 2, lnoise_range = c(5, 12),
    diameter_range = c(16, 22), threshold_range = c(5, 15, 30), 
    verbose = FALSE, plot = TRUE)
```

**Note**: the `getOptimizedParams()` is a *getter* function to obtain the values
of each optimized parameter.

```{r echo=TRUE, include=TRUE, eval=TRUE}
# obtain optimized params
getOptimizedParams(x1)$auto_params
```


#### Tracking Cell Movements

The central step of *Module 1* is tracking cell movements across all frames of a
multi-stack image (where each stack was acquired at a different time). This
operation is carried out via the `CellTracker()` function, which performs two
tasks: *i)* identify all cells in each frame of the image; *ii)* map cells
across all image frames, identify cell movements and return cell tracks.
This operation supports parallelization. This function requires three
parameters to be set: `lnoise`, `diameter`, and `threshold`. These parameters
can be set manually or automatically:

- rely on the optimized params estimated using `OptimizeParams()`

- rely on the optimized params estimated for a different `trackedCells` object;
using `OptimizeParams()`; see the `import_optiParam_from` argument

- the user can manually specify the parameter values; note that user-specified
parameters will overwrite automatically-optimized values

**Note 1**: the user may request to visualize a plot for each frame being
processed. The output plot shows cells that were detected for each combination
of parameter values.

**Note 2**: it is possible to only include cells that were detected in at least
a minimum number of frames by setting the `min_frames_per_cell` argument. If so,
cells detected in a small number of frames will be removed from the output.


```{r echo=TRUE, include=TRUE, eval=TRUE, fig.keep='last'}
# Track cell movements using optimized params
x1 <- CellTracker(
    tc_obj = x1, min_frames_per_cell = 3, threads = 1, verbose = TRUE)

# Track cell movements using params from a different object
x2 <- CellTracker(
    ThreeConditions$ctrl02, import_optiParam_from = x1,
    min_frames_per_cell = 3, threads = 2)
```

```{r fig.height=4, fig.width=4, fig.align='center'}
# Track cell movements using CUSTOM params, show plots
x3 <- CellTracker(
    tc_obj = ThreeConditions$drug01,
    lnoise = 5, diameter = 22, threshold = 6,
    threads = 2, maxDisp = 10,
    show_plots = TRUE)
```


It is possible to retrieve the output data.frame including information about
cell movements (cell tracks) using the `getTracks()` getter function.

```{r echo=TRUE, include=TRUE, eval=TRUE, results='asis'}
# Get tracks and show header
trk1 <- cellmigRation::getTracks(x1)
head(trk1) %>% kable() %>% kable_styling(bootstrap_options = 'striped')
```


#### Basic migration stats

For compatibility and portability reasons, Module 1 includes a function to
compute the same basic metrics/stats as in the *FastTracks* Matlab software by
Brian DuChez. This step is performed via the `ComputeTracksStats()` function.
The results can be extracted from a `trackedCells` object via dedicated getter
functions: `getPopulationStats()` and `getCellsStats()`. Note however that
more advanced stats are computed using functions included in the second module
of `cellmigRation`.

```{r echo=TRUE, include=TRUE, eval=TRUE}
# Basic migration stats can be computed similar to the fastTracks software
x1 <- ComputeTracksStats(
    x1, time_between_frames = 10, resolution_pixel_per_micron = 1.24)
x2 <- ComputeTracksStats(
    x2, time_between_frames = 10, resolution_pixel_per_micron = 1.24)
x3 <- ComputeTracksStats(
    x3, time_between_frames = 10, resolution_pixel_per_micron = 1.24)

# Fetch population stats and attach a column with a sample label
stats.x1 <- cellmigRation::getCellsStats(x1) %>%
    mutate(Condition = "CTRL1")
stats.x2 <- cellmigRation::getCellsStats(x2) %>%
    mutate(Condition = "CTRL2")
stats.x3 <- cellmigRation::getCellsStats(x3) %>%
    mutate(Condition = "DRUG1")
```

```{r echo=TRUE, include=TRUE, eval=TRUE, results='asis'}
stats.x1 %>%
    dplyr::select(
        c("Condition", "Cell_Number", "Speed", "Distance", "Frames")) %>%
    kable() %>% kable_styling(bootstrap_options = 'striped')
```

```{r fig.height=4.5, fig.width=4.9, fig.align='center'}
# Run a simple Speed test
sp.df <- rbind(
    stats.x1 %>% dplyr::select(c("Condition", "Speed")),
    stats.x2 %>% dplyr::select(c("Condition", "Speed")),
    stats.x3 %>% dplyr::select(c("Condition", "Speed"))
)

vp1 <- ggplot(sp.df, aes(x=Condition, y = Speed, fill = Condition)) +
    geom_violin(trim = FALSE) +
    scale_fill_manual(values = c("#b8e186", "#86e1b7", "#b54eb4")) +
    geom_boxplot(width = 0.12, fill = "#d9d9d9")

print(vp1)
```

```{r echo=TRUE, include=TRUE, eval=TRUE, results='markup'}
# Run a t-test:
sp.lst <- with( sp.df, split(Speed, f = Condition))
t.test(sp.lst$CTRL1, sp.lst$DRUG1, paired = FALSE, var.equal = FALSE)
```

#### Basic Visualizations

Two basic visualization functions are included in Module 1, and allow
visualization of cells detected in a frame of interest, and tracks originating
at a frame of interest. These functions are included in Module 1 (and not
Module 2) since they take a `trackedCells`-class object as input.

```{r fig.height=4, fig.width=4, fig.align='center'}
# Visualize cells in a frame of interest
cellmigRation::VisualizeStackCentroids(x1, stack = 1)
```

```{r echo=TRUE, include=TRUE, eval=TRUE, fig.height=4.6}
# Visualize tracks of cells originating at a frame of interest
par(mfrow = c(1, 3))
cellmigRation::visualizeCellTracks(x1, stack = 1, main = "tracks from CTRL1")
cellmigRation::visualizeCellTracks(x2, stack = 1, main = "tracks from CTRL2")
cellmigRation::visualizeCellTracks(x3, stack = 1, main = "tracks from DRUG1")
```


#### Aggregate Cell Tracks

Cell tracks from multiple TIFF images can be aggregated together. All tracks
form the different experiments/images are returned in a large data.frame. A new
unique ID is assigned to specifically identify each cell track from each
image/experiment. Different `trackedCells` objects can be merged together based
on the corresponding *TIFF filename* (default), or one of the meta-information
included in the object(s).

**Note 1**: the data.frame returned by `aggregateTrackedCells()` has a structure
that aligns to the output of the `getTracks()` function when the `attach_meta`
argument is set to TRUE.

**Note 2**: the data.frame returned by `aggregateTrackedCells()` (or by
`getTracks()` with the `attach_meta` argument set to TRUE) is the input of
the `CellMig()` function, and is the first step of Module 2.

**Note 3**: it is recommended to aggregate experiments/tiff files corresponding
to the same condition (as shown below: for example, all replicates of the
control cells) However, it is also possible to mix and match multiple
treatments/timepoints/conditions, and filter the desired tracks right before
running the `CellMig()` step (not shown).

```{r echo=TRUE, include=TRUE, eval=TRUE, results='asis'}
# aggregate tracks together
all.ctrl <- aggregateTrackedCells(x1, x2, meta_id_field = "tiff_file")

# Show header
all.ctrl[seq_len(10), seq_len(6)] %>%
    kable() %>% kable_styling(bootstrap_options = 'striped')
```


```{r eval = TRUE, echo=FALSE, results='markup', include=TRUE}
# Table tiff_filename vs. condition
with(all.ctrl, table(condition, tiff_file))
```

```{r eval = TRUE, echo = TRUE, results='markup', include=TRUE}
# Prepare second input of Module 2
all.drug <- getTracks(tc_obj = x3, attach_meta = TRUE)
```


### Module 2

The second module of `cellmigRation` is aimed at computing advanced stats and
building 2D, 3D, and interactive visualizations based on the cell tracks
computed in Module 1.

#### Import and Pre-process Cell Tracks

The first step entails the generation of a `CellMig`-class object (S4 class)
to store cell tracks data, and all output resulting from running Module 2
functions. After importing data into a `CellMig`-class object, tracks are
processed according to the experiment type (random migration in a plate vs.
scratch-wound healing assay).

**Note 1**: the arguments passed to the `CellMig()` function are:

- **trajdata** a data.frame, the output from the previous module

- **expName** a string, this is the name of the experiment


**Note 1**: the user is allowed to name the analysis; here we select a name
that will be used as a prefix in the name of plots and tables.

**Note 2**: For Random Migration assays, the `rmPreProcessing()` function is
used for preprocessing; if a Scratch Wound Healing Assay was performed, the
`wsaPreProcessing()` function shall be used instead.


```{r eval = TRUE, echo = TRUE, results='markup', include=TRUE}
rmTD <- CellMig(trajdata = all.ctrl)
rmTD <- setExpName(rmTD, "Control")

# Preprocessing the data
rmTD <- rmPreProcessing(rmTD, PixelSize=1.24, TimeInterval=10, FrameN=3)
```

#### Plotting tracks (2D and 3D)

Multiple plotting functions allow the user to generate 2D or 3D charts and
plots showing the movements of all cells, or part of the cells in the
experiment.

```{r eval = TRUE, echo = TRUE, include=TRUE, fig.width=5}
# Plotting tracks (2D and 3D)
plotAllTracks(rmTD, Type="l", FixedField=FALSE, export=FALSE)
```

```{r eval = TRUE, echo = TRUE, include=TRUE, fig.keep='last', fig.width=5}
# Plotting the trajectory data of sample of cells (selected randomly)
# in one figure
plotSampleTracks(
    rmTD, Type="l", FixedField=FALSE, celNum=2, export = FALSE)
```

#### 3D Plots

The following functions are meant to be run in an interactive fashion:

- `plot3DAllTracks(rmTD, VS=2, size=5)`

- `plot3DTracks(rmTD, cells=1:10, size = 8)`


#### Deep Trajectory Analysis

The deep trajectory analysis includes a series of tools to examine the
following metrics:

- Persistence and Speed: `PerAndSpeed()` function

- Directionality: `DiRatio()` function

- Mean Square Displacement: `MSD()` function

- Direction AutoCorrelation: `DiAutoCor()` function

- Velocity AutoCorrelation: `VeAutoCor()` function

These steps are meant to be run on larger datasets, including a larger number 
of cells. Here, we only show an example of how to run a *DiRatio* analysis, 
an *MSD* analysis and *Velocity autocorrelation*. 

**For more examples about Deep Trajectory Analysis, please visit:**
<https://www.data-pulse.com/projects/2020/cellmigRation/cellmigRation_v01.html>


**Directionality Analysis**. This analysis is performed via the `DiRatio()` 
function. Results are saved in a *CSV* file. Plots can be generated using the 
`DiRatioPlot()` function. Plots are saved in a newly created folder with the 
following extension: `-DR_Results`.


```{r echo=TRUE, include=TRUE, eval=TRUE}
## Directionality
srmTD <- DiRatio(rmTD, export=TRUE)
DiRatioPlot(srmTD, export=TRUE)
```


```{r, echo=FALSE, out.width="50%", fig.cap="Controldirectionality"}
knitr::include_graphics(
    "Control-DR_Results/Controldirectionality ratio for all cells.jpg")
```



**Mean Square Displacement**.  The MSD function automatically computes the mean
square displacements across several sequential time intervals. MSD parameters
are used to assess the area explored by cells over time. Usually, both the
`sLAG` and `ffLAG` arguments are recommended to be set to 0.25 but since here
we have only few frames per image, we will set it to 0.5.

```{r echo=TRUE, include=TRUE, eval=TRUE}
rmTD<-MSD(object = rmTD, sLAG=0.5, ffLAG=0.5, export=TRUE)
```



**Velocity AutoCorrelation**. The `VeAutoCor()` function automatically computes
the changes in both speed and direction across several sequantial time
intervals. Usually the `sLAG` is recommended to be set to 0.25 but since here
we have just few frames, we will set it to 0.5.

```{r echo=TRUE, include=TRUE, results='asis', fig.keep='last', fig.width=5}
rmTD <- VeAutoCor(
    rmTD, TimeInterval=10, sLAG=0.5, sPLOT=TRUE,
    aPLOT=TRUE, export=FALSE)
```

#### Final Results

The `FinRes()` function automatically generates a data frame that contains all
the results with or without the a correlation table.

```{r echo=TRUE, include=TRUE, eval=TRUE, message=FALSE, warning=FALSE}
rmTD <-FinRes(rmTD, ParCor=TRUE, export=FALSE)
```

Below, the first 5 columns of the output data.frame are shown.

```{r echo=FALSE, include=TRUE, results='asis'}
head(getCellMigSlot(rmTD, "results"), 5) %>% 
    kable() %>% kable_styling(bootstrap_options = 'striped')
```

#### Principal Component Analysis (PCA) and Cell Clustering

The `CellMigPCA()` function automatically generates Principal Component Analysis
based on a set of parameters selected by the user.
The `CellMigPCAclust()` function automatically generates clusters based on the
Principal Component Analysis. This analysis is supposed to be run in an 
interactive session via the `CellMigPCA()` function.


### Session & Environment

```{r include=FALSE}
Tim1 <- Sys.time()
TimDiff <- as.numeric(difftime(time1 = Tim1, time2 = Tim0, units = "mins"))
TimDiff <- format(round(TimDiff, digits = 2), nsmall = 2)
```

- **Execution time**: vignette built in: `r TimDiff` minutes.

- **Session Info**: shown below.

```{r echo=FALSE, eval=TRUE, results='markup'}
sessionInfo()
```

**Success!** For questions about `cellmigRation`, don't hesitate to email the
authors or the maintainer.

